\textbf{Regressions And Stuff}

### 2016-09-30 paper2/paper.tex
more problem 1 stuff

\documentclass[letterpaper, twocolumn,10pt]{article}
\usepackage{epsfig,endnotes,amsmath,usenix}

@@

In every step of the itreration, we update our guess as following:
\begin{align*}
\mathbf{x}_{n+1} = \mathbf{x} - \lambda \nabla f(\mathbf{x}_{n})
\end{align*}
where $\lambda$ is the \emph{step size} of the iteration. The algorith terminates upon the following convergence condition: 
\begin{align*}
|f(\mathbf{x}_{n+1}) - f(\mathbf{x}_{n})| < \delta
\end{align*}
where $\delta$ is the convergence \emph{threshold}. Upon convergence, the algorithm returns a final guess of $\mathbf{x}_{\text{opt}} = \mathbf{x}_{n+1}$.

In this section, we compare two methods of gradient descent: \emph{batch gradient descent}, which computes a gradient over all samples for each iteration, and \emph{stochastic gradient descent}, which uses pointwise gradients instead.

To analyze the perforamnce of the different algorithms, we use the \emph{Gaussian function}:
\begin{align*}
f(x) = - \dfrac{1}{\sqrt{(2\pi)^n |\sigma|}} \exp \left[-\dfrac{1}{2} (x - u)^T \sigma^{-1} (x-u)\right]
\end{align*}
In addition, we use the \emph{quadratic bowl function}:
\begin{align*}
f(x) = \dfrac{1}{2} x^T A x - x^T b 
\end{align*}
Finally, we use the \emph{least squares error} function:
\begin{align*}
J(\theta) = |X\theta - y|^2
\end{align*}

We analyze the affects of step size ($\lambda$), threshold ($\delta$), and initial guess $\mathbf{x}_{0}$ on the performance and accuracy of the gradient descent algorithm.

@@

\subsection{Stochastic Gradient Descent}
\subsection{Effect of Finite Gradient}

>>>>

\documentclass{paper}
\usepackage{fullpage, amsmath, graphicx}

@@

In every step of the itreration, we update our guess as following:
\begin{align*}
\mathbf{x}_{n+1} = \mathbf{x} - \lambda \nabla f(\mathbf{x}_{n})
\end{align*}
where $\lambda$ is the \emph{step size} of the iteration. The algorithm terminates upon the following convergence condition: 
\begin{align*}
|f(\mathbf{x}_{n+1}) - f(\mathbf{x}_{n})| < \delta
\end{align*}
where $\delta$ is the convergence \emph{threshold}. Upon convergence, the algorithm returns a final guess of $\mathbf{x}_{\text{opt}} = \mathbf{x}_{n+1}$.

In this section, we compare two methods of gradient descent: \emph{batch gradient descent}, which computes a gradient over all samples for each iteration, and \emph{stochastic gradient descent}, which instead uses pointwise gradients.

To analyze the perforamnce of the different algorithms, we use the \emph{Gaussian function}:
\begin{align*}
f(x) = - \dfrac{1}{\sqrt{(2\pi)^n |\Sigma|}} \exp \left[-\dfrac{1}{2} (x - u)^T \Sigma^{-1} (x-u)\right]
\end{align*}
In addition, we use the \emph{quadratic bowl function}:
\begin{align*}
f(x) = \dfrac{1}{2} x^T A x - x^T b 
\end{align*}
Finally, we use the \emph{least squares error} function:
\begin{align*}
J(\theta) = |X\theta - y|^2
\end{align*}

@@

In batch gradient descent, for every iteration we calculate the gradient across the entire sample set. In this section, we analyze the affects of step size ($\lambda$), threshold ($\delta$), and initial guess $\mathbf{x}_{0}$ on the performance and accuracy of the gradient descent algorithm.

First, we note that step size has a significant impact on the convergence rate of the algorithm. We run gradient descent on an arbitrary quadratic bowl function. As we increase the value of $\lambda$ from $0.0001$ to $0.001$, the magnitude of the gradient decreases more quickly, and the algorithm converges faster. Note that in this figure, all of the gradient descents were run with a fixed size of 2000 iterations to illustrate the effects on the gradient over time.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.5\textwidth]{../images/quadratic_steps_vs_mag}
  \caption{As the step size increases, the algorithm converges quicker.}
\end{figure}

In addition, the step size and threshold control the accuracy of the gradient descent. As step size decreases, we expect a more accurate result because the function of the initial guess over time is smoother. Similarly, as the threshold decreases, we expect a more accurate result because the algorithm needs to run for longer before reaching a stable equilibrium. To verify these claims, we run gradient descent using a two-dimensional Gaussian with mean $(10, 10)$. As we vary $\lambda$ and $\delta$, we plot the least-squares error between the returned value and the actual optimum of $(10, 10)$ (see Figure 2). Indeed, the error decreases as $\lambda$ and $\delta$ decrease.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.5\textwidth]{../images/lsq_error_gradient}
  \caption{As $\lambda$ and $\delta$ decrease, error decreases.}
\end{figure}

Finally, we observe the affect of the initial case on the rate of convergence. When running gradient descent on a two-dimensional Gaussian with mean $(10, 10)$, one would expect that as the intial guess appraoches the optimal value, the convergence rate increases. Indeed, running the algorithm with $\lambda = 1000$ and $\delta = 10^{-11}$, we notice that the rate of convergence is much higher as the initial guess approaches the actual value (see Figure 3).

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.5\textwidth]{../images/guess_vs_mags_2}
  \caption{As $\mathbf{x}_0$ appraoches $(10, 10)$, the rate of convergence increases.}
\end{figure}

\subsection{Stochastic Gradient Descent}

\subsection{Effect of Finite Gradient}

When the objective function doesn't have a closed form solution, one can approximate the gradient using a central finite difference. In other words, we have 
\begin{align*}
\nabla f(x) \approx \nabla_h[f](x) \dfrac{f(x+h/2) - f(x-h/2)}{h}
\end{align*}
Therefore, we only need the objective function to approximate the gradient. Note that as $h$ approaches zero, this approaches the actual value of the gradient; and has $h$ approaches infinity this value diverges. Figure 4 illustrates the impact of $h$ on the accuracy of the gradient: for small enough $h$, the approximation is accurate to several orders of magnitude.

\begin{figure}[ht!]
  \centering
    \begin{tabular}{c | c}
    $h$  & $\nabla_h[f](x)$ \\	\hline
	0.01 & $-3.1704 \cdot 10^{-7}$ \\
	0.1 & $-3.1704 \cdot 10^{-7}$ \\
	1.0 & $-3.1696 \cdot 10^{-7}$ \\
	10.0 & $-3.0923 \cdot 10^{-7}$ \\
	100.0 & $-2.6198 \cdot 10^{-8}$ \\
	1000.0 & $\approx 0$ \\
    \end{tabular}
  \caption{As $h$ increases, the accuracy of the central finite difference decreases. The actual gradient value is $-3.1704 \cdot 10^{-7}$.}
\end{figure}
